19Q1

### 2019-04-30 19Q1.Rmd
update rmd, start on re-making Fig1

```{r}
library(cowplot)
library(crunch)
library(ggpubr)
library(ggrepel)
library(ggsignif)
library(kableExtra)
library(naniar)
library(openxlsx)
library(reshape2)
library(tidyverse)
library(VennDiagram)

@@

```{r functions}
unfactorize <- function(df){
  for(i in which(sapply(df, class) == "factor")) {
    df[[i]] = as.character(df[[i]])
  }
  return(df)
}

MergeOmicsG2P <- function(database, outfile) {
  # Merge mutation status, CN, and GE omics data with drugscreen results and metadata
  if(database %in% c("CCLE", "CTRP", "GDSC")) {
    filt <- auc %>% filter(Database == database) %>% select(Cell_line, Histology, TCGA_classification, CCLE_Name, Broad_ID, Drug, AUC_IC50)
    print(paste0("Processing ", database, " data."))
    colnames(filt)[colnames(filt) == "AUC_IC50"] <- "Score"

# Data

@@

```{r}
g2p <- read.delim("./data_munging/g2p_full_level_A_dataset_NEW_DEC_2_2018_withGroupedLineages.csv", sep = "\t", header = TRUE)
g2p[, 1:2] <- NULL
g2p <- merge(g2p, lineage_converter, by = "grouped_general_doid", all.x = TRUE)

g2p$Drug <- tolower(g2p$Drug)
g2p[g2p == -1] <- NA
g2p$Ref[g2p$Ref == "None"] <- NA
g2p$Alt[g2p$Alt == "None"] <- NA
g2p$Drug_Gene <- ifelse(is.na(g2p$Drug) | is.na(g2p$Gene), NA, paste0(g2p$Drug, "_", g2p$Gene))
g2p$Drug_Gene_Lin <- ifelse(is.na(g2p$Drug) | is.na(g2p$Gene) | is.na(g2p$group_general_lineage_name), NA, paste0(g2p$Drug, "_", g2p$Gene, "_", g2p$group_general_lineage_name))
g2p$position <- paste0("g.chr", g2p$Chromosome, ":", g2p$End, g2p$Ref, ">", g2p$Alt)
g2p$position <- ifelse(grepl("NA", g2p$position), NA, g2p$position)
g2p_filt <- subset(g2p, grepl("^CID", g2p$DrugID))

g2p_dgls <-  unique(g2p_filt$Drug_Gene_Lin[!is.na(g2p_filt$Drug_Gene_Lin)])
g2p_druggenes <- unique(g2p_filt$Drug_Gene[!is.na(g2p_filt$Drug_Gene)])
g2p_drugs <- unique(sapply(strsplit(g2p_dgls, "_", fixed = TRUE), function(x) x[1]))
g2p_genes <- unique(sapply(strsplit(g2p_dgls, "_", fixed = TRUE), function(x) x[2]))
g2p_lins <- unique(sapply(strsplit(g2p_dgls, "_", fixed = TRUE), function(x) x[3]))
g2p_druggenes
g2p_drugs
g2p_genes
g2p_lins

@@

```{r, eval=FALSE}
# levels(g2p_filt$Phenotype.Description)
unique(g2p_filt$group_general_lineage_name)
levels(g2p_filt$group_general_lineage_name)

g2p_nonCID <- subset(g2p, !grepl("^CID", g2p$DrugID))
summ <- g2p_nonCID %>% group_by(Drug, DrugID) %>% tally() %>% summarize(DrugID = DrugID, n = n, percent = formatC(n / 402 * 100, digits = 6, format = "f"))
summ
```

@@

@@

### Merge omics

@@

`r nrow(filter(test, n != 1))`/`r nrow(test)` gene-cell line combinations have more than one mutation in the MAF file. This is corrected in the point mutation filtering done below.

**NEED TO RUN CTRP OVER WEEKEND**
```{r, eval=FALSE}
# Load in Pozdeyev et al. (2016) data
ccle_all <- read.delim("./../htscreens_giant_files/pozdeyev_data/CCLE_drug_response_metrics.csv.gz", sep = ",")
ctrp_all <- read.delim("./../htscreens_giant_files/pozdeyev_data/CTRP_drug_response_metrics.csv.gz", sep = ",")
gdsc_all <- read.delim("./../htscreens_giant_files/pozdeyev_data/GDSC_drug_response_metrics.csv.gz", sep = ",")

@@

### Summaries

@@

### Merge CRISPR data and annotations

@@

# Wilcoxon tests

@@

# Spearman correlations

@@

Spearman correlations for gene expression and copy number:
```{r fig.height=6, fig.width=12}
crispr_spear_ge <- transform(crispr_g2p %>% group_by(Gene) %>% summarize(p = cor.test(y = Score, x = TPM, method = "spearman", use = "complete.obs")$p.value, p_log10 = -log10(p), corr = cor.test(y = Score, x = TPM, method = "spearman", use = "complete.obs")$estimate, Metric = "Gene Expression"), Gene = reorder(Gene, corr))
crispr_spear_cn <- transform(crispr_g2p %>% group_by(Gene) %>% summarize(p = cor.test(y = Score, x = CN, method = "spearman", use = "complete.obs")$p.value, p_log10 = -log10(p), corr = cor.test(y = Score, x = CN, method = "spearman", use = "complete.obs")$estimate, Metric = "Copy Number"), Gene = reorder(Gene, corr))

spearman_KRAS <- rbind(filter(crispr_g2p, Gene == "KRAS") %>% summarize(p = cor.test(y = Score, x = TPM, method = "spearman", use = "complete.obs")$p.value, corr = cor.test(y = Score, x = TPM, method = "spearman", use = "complete.obs")$estimate, Metric = "Gene Expression"), filter(crispr_g2p, Gene == "KRAS") %>% summarize(p = cor.test(y = Score, x = CN, method = "spearman", use = "complete.obs")$p.value, corr = cor.test(y = Score, x = CN, method = "spearman", use = "complete.obs")$estimate, Metric = "Copy Number"))

@@

Kolmogorov-Smirnov tests comparing DGLs that are in a G2P DGL and not.
```{r, eval=TRUE, fig.height=6, fig.width=12}
# g2p_corr_all <- G2PCorrTable(ccle_ptmuts_df = ccle_ptmuts, ctrp_ptmuts_df = ctrp_ptmuts, gdsc_ptmuts_df = gdsc_ptmuts)
# saveRDS(g2p_corr_all, "./data_munging/rds/g2p_corr_all.rds")
g2p_corr_all <- readRDS("./data_munging/rds/g2p_corr_all.rds")
g2p_mut_stats <- G2PCorrStatsTables(corr_df = g2p_corr_all, table_type = "mutation")
g2p_corr_stats <- G2PCorrStatsTables(corr_df = g2p_corr_all, table_type = "KS")

g2p_corr_4plot <- g2p_corr_all
g2p_corr_4plot$p <- NULL
g2p_corr_4plot$Mut_Signif <- NULL
g2p_corr_4plot <- melt(g2p_corr_4plot, id.vars = colnames(g2p_corr_4plot)[1:6], measure.vars = colnames(g2p_corr_4plot)[7:8], value.name = "Correlation")
g2p_corr_4plot <- g2p_corr_4plot %>% separate(col = variable, into = c("Correlation_Method", "corr", "Metric"), sep = "_", remove = FALSE)
g2p_corr_4plot$Metric <- ifelse(g2p_corr_4plot$Metric == "GE", "Gene Expression", "Copy Number")

g2p_ks_res_plot_spearman <- ggplot(data = filter(g2p_corr_4plot, Correlation_Method == "Spearman")) +
  facet_grid(Metric ~ Dataset, labeller = labeller(yfacet = c('CN' = "Copy Number", 'GE' = "Gene Expression"))) +
  geom_violin(aes(y = Correlation, x = G2P_Drug_Gene_Lin, fill = Dataset, color = Dataset, alpha = G2P_Drug_Gene_Lin)) +
  geom_boxplot(width = 0.075, aes(y = Correlation, x = G2P_Drug_Gene_Lin, color = Dataset)) +
  geom_jitter(width = 0.3, alpha = 0.8, aes(y = Correlation, x = G2P_Drug_Gene_Lin, color = Dataset)) +
  theme(legend.position = "none") +
  scale_alpha_manual(values = c("Yes" = 0.6, "No" = 0.2)) +
  scale_fill_manual(values = c("CCLE" = "springgreen4", "CTRP" = "steelblue4", "GDSC" = "turquoise4")) +
  scale_color_manual(values = c("CCLE" = "springgreen4", "CTRP" = "steelblue4", "GDSC" = "turquoise4")) +
  coord_cartesian(ylim = c(-1, 1)) +
  geom_text(data = filter(g2p_corr_stats, Correlation_Method == "Spearman"), aes(x = 1.5, y = 0.8, label = paste0("D = ", formatC(D_stat, digits = 2, format = "f"), ", p = ", ifelse(pval > 0.1, round(pval, 2), round(pval, 3)), "\n", ifelse(is.na(signif), "", signif)))) +
  labs(x = "Is the lineage in a G2P DGL?", y = "Spearman Correlation Coefficient")
g2p_ks_res_plot_spearman

# ggsave("./plots/manuscript/g2p_spearman_ks_res_plot.pdf", g2p_ks_res_plot_spearman, device = "pdf", dpi = 500, width = 12, height = 6, units = "in")
```

>>>>

```{r}
library(circlize)
library(ComplexHeatmap)
library(cowplot)
library(crunch)
library(ggpubr)
library(ggrepel)
library(ggsignif)
library(kableExtra)
library(naniar)
library(openxlsx)
library(reshape2)
library(tidyverse)
library(VennDiagram)

@@

```{r general_functions}
unfactorize <- function(df){
  for(i in which(sapply(df, class) == "factor")) {
    df[[i]] = as.character(df[[i]])
  }
  return(df)
}
```

# Import data

@@

```{r}
g2p_all <- read.delim("./data_munging/g2p_full_level_A_dataset_NEW_DEC_2_2018_withGroupedLineages.csv", sep = "\t", header = TRUE)
g2p_all[, 1:2] <- NULL
g2p_all$Drug <- tolower(g2p_all$Drug)
g2p_all[g2p_all == -1] <- NA
g2p_all$Ref[g2p_all$Ref == "None"] <- NA
g2p_all$Alt[g2p_all$Alt == "None"] <- NA
g2p_all <- subset(g2p_all, grepl("^CID", g2p_all$DrugID))
g2p <- merge(g2p_all, lineage_converter, by = "grouped_general_doid", all.x = TRUE)

g2p$Drug_Gene <- ifelse(is.na(g2p$Drug) | is.na(g2p$Gene), NA, paste0(g2p$Drug, "_", g2p$Gene))
g2p$Drug_Gene_Lin <- ifelse(is.na(g2p$Drug) | is.na(g2p$Gene) | is.na(g2p$group_general_lineage_name), NA, paste0(g2p$Drug, "_", g2p$Gene, "_", g2p$group_general_lineage_name))
g2p$position <- paste0("g.chr", g2p$Chromosome, ":", g2p$End, g2p$Ref, ">", g2p$Alt)
g2p$position <- ifelse(grepl("NA", g2p$position), NA, g2p$position)

g2p_dgls <-  unique(g2p$Drug_Gene_Lin[!is.na(g2p$Drug_Gene_Lin)])
g2p_druggenes <- unique(g2p$Drug_Gene[!is.na(g2p$Drug_Gene)])
g2p_drugs <- unique(sapply(strsplit(g2p_dgls, "_", fixed = TRUE), function(x) x[1]))
g2p_genes <- unique(sapply(strsplit(g2p_dgls, "_", fixed = TRUE), function(x) x[2]))
g2p_lins <- unique(sapply(strsplit(g2p_dgls, "_", fixed = TRUE), function(x) x[3]))
g2p_druggenes
g2p_drugs
g2p_genes
g2p_lins

@@

```{r, eval=FALSE}
unique(g2p$group_general_lineage_name)
levels(g2p$group_general_lineage_name)

g2p_nonCID <- subset(g2p_all, !grepl("^CID", g2p_all$DrugID))
summ <- g2p_nonCID %>% group_by(Drug, DrugID) %>% tally() %>% summarize(DrugID = DrugID, n = n, percent = formatC(n / 402 * 100, digits = 6, format = "f"))
summ
```

@@

# Merge omics

```{r omics_functions}
MergeOmicsG2P <- function(database, outfile) {
  # Merge mutation status, CN, and GE omics data with drugscreen results and metadata
  if(database %in% c("CCLE", "CTRP", "GDSC")) {
    filt <- auc %>% filter(Database == database) %>% select(Cell_line, Histology, TCGA_classification, CCLE_Name, Broad_ID, Drug, AUC_IC50)
    print(paste0("Processing ", database, " data."))
    colnames(filt)[colnames(filt) == "AUC_IC50"] <- "Score"

@@

@@

`r nrow(filter(test, n != 1))`/`r nrow(test)` gene-cell line combinations have more than one mutation in the MAF file. This is corrected in the point mutation filtering done below.
```{r, eval=FALSE}
# Load in Pozdeyev et al. (2016) data
ccle_all <- read.delim("./../htscreens_giant_files/pozdeyev_data/CCLE_drug_response_metrics.csv.gz", sep = ",")
ctrp_all <- read.delim("./../htscreens_giant_files/pozdeyev_data/CTRP_drug_response_metrics.csv.gz", sep = ",")
gdsc_all <- read.delim("./../htscreens_giant_files/pozdeyev_data/GDSC_drug_response_metrics.csv.gz", sep = ",")

@@

@@

@@

# G2P summary figures

```{r fig.height=6, fig.width=12}
g2p_dg_counts <- g2p_all %>% count(Drug, , Gene) %>% filter(!is.na(Drug), !is.na(Gene))
g2p_dg_counts <- dcast(g2p_dg_counts, Gene ~ Drug, value.var = "n", fun.aggregate = mean, fill = 0) %>% `rownames<-`(.[,1]) %>% select(-Gene)
g2p_dg_counts <- g2p_dg_counts[order(rowSums(-g2p_dg_counts)), order(colSums(-g2p_dg_counts))]
g2p_dg_counts[g2p_dg_counts == 0] <- NA

# Define color scale
col_fun <- colorRamp2(c(0, 103), c("mistyrose", "purple4"))
# Legend properties list
ht_legend_list <- list(legend_direction = "horizontal", title_position = "lefttop", legend_width = unit(6, "in"), border = "black")
ht <- Heatmap(g2p_dg_counts, 
              name = "Frequency",
              col = col_fun,
              na_col = "white",
              rect_gp = gpar(col = "gray88"),
              heatmap_legend_param = ht_legend_list,
              cluster_rows = FALSE, 
              cluster_columns = FALSE, 
              row_names_side = "left",
              row_title = "Genes",
              column_title = "Drugs", 
              column_title_side = "bottom")

# png(filename = out_filepath, width = 7.5, height = 8, units = "in", res = 500)
draw(ht, heatmap_legend_side = "top")
# dev.off()
```

# Wilcoxon tests (mut)

@@

# Spearman (CN, GE)

@@

Spearman correlations for gene expression and copy number:
```{r}
crispr_spear_ge <- transform(crispr_g2p %>% group_by(Gene) %>% summarize(p = cor.test(y = Score, x = TPM, method = "spearman", use = "complete.obs")$p.value, p_log10 = -log10(p), corr = cor.test(y = Score, x = TPM, method = "spearman", use = "complete.obs")$estimate, Metric = "Gene Expression"), Gene = reorder(Gene, corr))
crispr_spear_cn <- transform(crispr_g2p %>% group_by(Gene) %>% summarize(p = cor.test(y = Score, x = CN, method = "spearman", use = "complete.obs")$p.value, p_log10 = -log10(p), corr = cor.test(y = Score, x = CN, method = "spearman", use = "complete.obs")$estimate, Metric = "Copy Number"), Gene = reorder(Gene, corr))
```

Create CRISPR two-column, KRAS-inset plot for CN and GE:
```{r fig.height=6, fig.width=12}
spearman_KRAS <- rbind(filter(crispr_g2p, Gene == "KRAS") %>% summarize(p = cor.test(y = Score, x = TPM, method = "spearman", use = "complete.obs")$p.value, corr = cor.test(y = Score, x = TPM, method = "spearman", use = "complete.obs")$estimate, Metric = "Gene Expression"), filter(crispr_g2p, Gene == "KRAS") %>% summarize(p = cor.test(y = Score, x = CN, method = "spearman", use = "complete.obs")$p.value, corr = cor.test(y = Score, x = CN, method = "spearman", use = "complete.obs")$estimate, Metric = "Copy Number"))

@@

Kolmogorov-Smirnov tests comparing DGLs that are in a G2P DGL and not.

**NEED TO FILTER FOR ONLY G2P DRUG-GENE COMBINATIONS PRIOR TO KS TEST**

```{r, eval=TRUE, fig.height=6, fig.width=12}
# g2p_corr_all <- G2PCorrTable(ccle_ptmuts_df = ccle_ptmuts, ctrp_ptmuts_df = ctrp_ptmuts, gdsc_ptmuts_df = gdsc_ptmuts)
# saveRDS(g2p_corr_all, "./data_munging/rds/g2p_corr_all.rds")
g2p_corr_all <- readRDS("./data_munging/rds/g2p_corr_all.rds")
g2p_mut_stats <- G2PCorrStatsTables(corr_df = g2p_corr_all, table_type = "mutation")

g2p_corr_4plot <- filter(g2p_corr_all, G2P_Drug_Gene == "Yes")
g2p_corr_stats <- G2PCorrStatsTables(corr_df = g2p_corr_4plot, table_type = "KS")
g2p_corr_4plot$p <- NULL
g2p_corr_4plot$Mut_Signif <- NULL
g2p_corr_4plot <- melt(g2p_corr_4plot, id.vars = colnames(g2p_corr_4plot)[1:6], measure.vars = colnames(g2p_corr_4plot)[7:8], value.name = "Correlation")
g2p_corr_4plot <- g2p_corr_4plot %>% separate(col = variable, into = c("Correlation_Method", "corr", "Metric"), sep = "_", remove = FALSE)
g2p_corr_4plot$Metric <- ifelse(g2p_corr_4plot$Metric == "GE", "Gene Expression", "Copy Number")

g2p_ks_res_plot_spearman <- ggplot(data = filter(g2p_corr_4plot, Correlation_Method == "Spearman")) +
  facet_grid(Metric ~ Dataset, labeller = labeller(yfacet = c('CN' = "Copy Number", 'GE' = "Gene Expression"))) +
  geom_violin(aes(y = Correlation, x = G2P_Drug_Gene_Lin, fill = Dataset, color = Dataset, alpha = G2P_Drug_Gene_Lin)) +
  geom_boxplot(width = 0.075, aes(y = Correlation, x = G2P_Drug_Gene_Lin, color = Dataset)) +
  geom_jitter(width = 0.3, alpha = 0.8, aes(y = Correlation, x = G2P_Drug_Gene_Lin, color = Dataset)) +
  theme(legend.position = "none") +
  scale_alpha_manual(values = c("Yes" = 0.6, "No" = 0.2)) +
  scale_fill_manual(values = c("CCLE" = "springgreen4", "CTRP" = "steelblue4", "GDSC" = "turquoise4")) +
  scale_color_manual(values = c("CCLE" = "springgreen4", "CTRP" = "steelblue4", "GDSC" = "turquoise4")) +
  coord_cartesian(ylim = c(-1, 1)) +
  geom_text(data = filter(g2p_corr_stats, Correlation_Method == "Spearman"),
            aes(x = 1.5, y = 0.8,
                label = paste0("D = ", formatC(D_stat, digits = 2, format = "f"),
                               ", p = ", ifelse(pval > 0.1, round(pval, 2), ifelse(pval < 0.001, formatC(pval, digits = 1, format = "e"), round(pval, 3))),
                               "\n", ifelse(is.na(signif), "", signif)))) +
  labs(x = "Is the lineage in a G2P DGL?",
       y = expression("Spearman Correlation Coefficient"~(r[s])))
g2p_ks_res_plot_spearman

ggsave("./plots/manuscript/g2p_spearman_ks_res_plot.png", g2p_ks_res_plot_spearman, device = "png", dpi = 500, width = 12, height = 6, units = "in")
```
